Marginalization of photometric redshift uncertainties

### 2020-07-22 Paper/paper.tex

      The derivative matrix can be seen in Fig.~\ref{fig:Tmat} for the first tomographic bin and its cross-correlations with all four bins in the HSC data, i.e. the tomographic pairs [1,1], [1,2], [1,3], and [1,4]. The rows of the matrix correspond to the $\ell$ multipole bandpowers available for each pair, and the columns to the 300 equally-spaced redshift samples of $N(z)$ between $z = 0$ and $z = 4$. We can notice that as we move to pairs with bins at higher redshifts, the peak of the matrix appears at different locations: namely, the high-$\ell$ high-$z$ regions of the matrix \da{I find this kind of confusing. If I look at the matrix, what I see is that there is a dominant column  which moves with the mean redshift of each bin, not so much about things moving in the $\ell$ direction in a coherent manner.}. In the $z$ direction, these coincide with the maxima of the corresponding redshift distribution.

      These $N(z)$s are shown as blue solid lines in Fig.~\ref{fig:Nzs} \da{We should describe either here or in the previous section what we define as the fiducial $N(z)$s. I guess, from the fact that the original and smoothed versions aren't so different, that the original ones are the sum of all photo-$z$ codes. Is that the case? Should we also show the COSMOS $N(z)$s and corresponding smoothed version? I'd be really curious to see what they look like. Right now it seems like the smoothing does absolutely nothing.}\bh{Resolved?}. The figure also shows the {\sl smoothed} distribution $\bar{N}(z)$s in solid red \da{maybe we should make it dashed so we can actually see the red line}\bh{Did that help?}, defined in Eq. \ref{eq:priorcov}. Our choice of smoothing prior has only a mild effect on the original redshift distribution, mostly removing sharp features at the edges of adjacent $N(z)$ measurements. Also shown in yellow is a realization of the $N(z)$ drawn from a multivariate Gaussian distribution with a mean given by the red line and a covariance given by ${\sl P}$ in Eq. \ref{eq:priorcov} \da{Is this true? If the smoothing prior has any relevance, then it should only allow "smooth" fluctuations. Also, the third bin is kind of suspicious in that all the fluctuations seem to be low.}

@@

      The angular power spectra, $C_\ell$, estimated for the same three sets of $N(z)$s are shown in Fig.~\ref{fig:Cls} using the same color scheme. The figure also shows the comparison between the exact prediction for the $N(z)$s (yellow line) and the linear prediction around the smoothed distribution using the derivative matrix, ${\sf T}$, (\textit{dotted light-blue line}). We find that the two are in reasonably good agreement given the statistical uncertainties in the data. \da{we could highlight that, better data should come with better $N(z)$ calibration samples so this statement is also true in the future.} It is worth noting that the agreement between exact prediction and linear model is noticeably worse for auto-correlations. As described in Appendix \ref{app:autos}, this is due to the fact that a single $N(z)$ enters the corresponding auto-correlation quadratically, making higher-order terms in Taylor expansion more relevant than in the case of cross-correlations. The $\chi^2$ values for all four curves with respect to the HSC data are shown in Table~\ref{tab:lin}. \da{I stopped here today} Comparing the $\chi^2$ values for the original and smoothed $N(z)$ distributions, we notice that they decrease for both the HSC covariance matrix and also for the marginalized one. As expected, perturbing the photometric distributions gets penalized, but the change in the $\chi^2$ value is much larger for the original HSC covariance, which indicates that the marginalized covariance is more lenient in allowing small (and expected) $N(z)$ fluctuations. The difference between the last two lines suggests that the Taylor expansion works sufficiently well, as the difference in $\chi^2$ is negligible. Finally, comparing the $\chi^2$ for the new and old covariances, we are led to conclude that the new covariance provides a better fit to the data. We note that we have fixed the values of all the shift and width parameter to be zero and set the 6 HOD parameters to their \textbf{fiducial} values of:
      \begin{itemize}
        \item $\mu_{\rm min} = 11.88$, $\mu_{\rm min,p} = -0.5$,
        \item $\mu_{\rm 0} = 5.7$, $\mu_{\rm 0,p} = 2.5$,
        \item $\mu_{\rm 1} = 13.08$, $\mu_{\rm 1,p} = 0.9$.
      \end{itemize}
      These HOD values are selected to match the 6 posterior means reported in Ref. \cite{1912.08209}. For the smoothed and perturbed $N(z)$ values only, we minimize the likelihood to find the best fit values for the 2 HOD parameters $\mu_1$ and $\mu_{1,p}$, obtaining 13.05 and 0.79, respectively. The ratio of  the original HSC precision matrix to the marginalized $N(z)$ one is  $1.15^{\nu_{\rm tot}}$, where $\nu_{\rm tot} = 94$ is the number of total degrees of freedom of the system (i.e. size of the $C_\ell$ vector for all cross correlations between the tomographic bins).

@@

>>>>

      The derivative matrix can be seen in Fig.~\ref{fig:Tmat} for the first tomographic bin and its cross-correlations with all four bins in the HSC data, i.e. the tomographic pairs [1,1], [1,2], [1,3], and [1,4]. The rows of the matrix correspond to the $\ell$ multipole bandpowers available for each pair, and the columns to the 300 equally-spaced redshift samples of $N(z)$ between $z = 0$ and $z = 4$. We can notice that as we move to pairs with bins at higher redshifts, the corresponding block of ${\sf T}$ peaks at a redshift column roughly corresponding to the maximum of the deepest distribution of the pair.

      These $N(z)$s are shown in Fig.~\ref{fig:Nzs}. We obtain the fiducial ``measured'' distributions $\hat{\vN}$, shown in solid blue in the Figure, using the COSMOS 30-band catalog re-weighted to account for the different colour space distribution of that sample using a nearest-neighbours approach as described in \da{cite}. The figure also shows the {\sl smoothed} distribution $\bar{N}$ in dashed red, defined in Eq. \ref{eq:priorcov}. Our choice of smoothing prior has only a mild effect on the original redshift distribution, mostly removing sharp features at the edges of adjacent $N(z)$ measurements. Also shown in solid yellow is a realization of $\vN$ drawn from a multivariate Gaussian distribution with a mean given by the red line and a covariance given by ${\sl P}$ in Eq. \ref{eq:priorcov} \da{Is this true? If the smoothing prior has any relevance, then it should only allow "smooth" fluctuations. Also, the third bin is kind of suspicious in that all the fluctuations seem to be low.}

@@

      The angular power spectra, $C_\ell$, estimated for the same three sets of $N(z)$s are shown in Fig.~\ref{fig:Cls} using the same color scheme. The figure also shows the comparison between the exact prediction for the $N(z)$s (yellow line) and the linear prediction around the smoothed distribution using the derivative matrix, ${\sf T}$, (\textit{dotted light-blue line}). We find that the two are in reasonably good agreement given the statistical uncertainties in the data. This may not be the case for future datasets with higher statistical power, although the expectation is that those data will be accompanied by larger spectroscopic samples \da{cite some of the Euclid efforts?} needed to reduce systematic photo-$z$ uncertainties. It is worth noting that the agreement between exact prediction and linear model is noticeably worse for auto-correlations. As described in Appendix \ref{app:autos}, this is due to the fact that a single $N(z)$ enters the corresponding auto-correlation quadratically, making higher-order terms in Taylor expansion more relevant than in the case of cross-correlations. The $\chi^2$ values for all four curves with respect to the HSC data are shown in Table~\ref{tab:lin}. \da{I stopped here today} Comparing the $\chi^2$ values for the original and smoothed $N(z)$ distributions, we notice that they decrease for both the HSC covariance matrix and also for the marginalized one. As expected, perturbing the photometric distributions gets penalized, but the change in the $\chi^2$ value is much larger for the original HSC covariance, which indicates that the marginalized covariance is more lenient in allowing small (and expected) $N(z)$ fluctuations. The difference between the last two lines suggests that the Taylor expansion works sufficiently well, as the difference in $\chi^2$ is negligible. Finally, comparing the $\chi^2$ for the new and old covariances, we are led to conclude that the new covariance provides a better fit to the data. We note that we have fixed the values of all the shift and width parameter to be zero and set the 6 HOD parameters to their \textbf{fiducial} values of:
      \begin{itemize}
        \item $\mu_{\rm min} = 11.88$, $\mu_{\rm min,p} = -0.5$,
        \item $\mu_{\rm 0} = 5.7$, $\mu_{\rm 0,p} = 2.5$,
        \item $\mu_{\rm 1} = 13.08$, $\mu_{\rm 1,p} = 0.9$.
      \end{itemize}
      These HOD values are selected to match the 6 posterior means reported in Ref. \cite{1912.08209}. For the smoothed and perturbed $N(z)$ values only, we minimize the likelihood to find the best fit values for the 2 HOD parameters $\mu_1$ and $\mu_{1,p}$, obtaining 13.05 and 0.79, respectively. The ratio of  the original HSC precision matrix to the marginalized $N(z)$ one is  $1.15^{\nu_{\rm tot}}$, where $\nu_{\rm tot} = 94$ is the number of total degrees of freedom of the system (i.e. size of the $C_\ell$ vector for all cross correlations between the tomographic bins).

@@

\section{Precision of the linear expansion for auto- and cross-correlations}\label{app:autos}
\da{Write this}
